---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip('activity.zip')
activity <- read.csv("activity.csv")
head(activity)
```


## What is mean total number of steps taken per day?
### Calculate total steps per day 
```{r}
stepsPerDay <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

### Plot them in a histogram
```{r}
hist(stepsPerDay)
```

### Get mean
```{r}
mean(stepsPerDay)
median(stepsPerDay)
```
Mean is 9354.23.
Median is 10395. 


## What is the average daily activity pattern?
### Create a time series plot
```{r}
avgDaily <- aggregate(x=list(avgSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
head(avgDaily)
plot(avgDaily, type="l")
```

### Get max steps
```{r}
rowOfMax <- which.max(avgDaily$avgSteps)
rowOfMax
maxInterval <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", avgDaily[rowOfMax,'interval'])
maxInterval
```
Interval is 8:30 - 8:35. 

## Imputing missing values
### Count missing values
```{r}
length(which(is.na(activity)))
```

### Strategy for replacing missing values
Replace NAs with mean.

### New dataset with filled missing values
```{r}
activityFilled <- activity
activityFilled[is.na(activityFilled[,1]), 1] <- mean(activityFilled[,1], na.rm = TRUE)

head(activityFilled)
```

### Histogram of daily steps (plus mean and median)
```{r}
dailySteps <- tapply(activityFilled$steps, activityFilled$date, sum)
hist(dailySteps)

mean(dailySteps)
median(dailySteps)
```


## Are there differences in activity patterns between weekdays and weekends?
### Create weekday, weekend variable
```{r}
activityFilled$dayType <-  ifelse(as.POSIXlt(activityFilled$date)$wday %in% c(0,6), 'weekend', 'weekday')

```

### Panel plot with times series plot for intervals and average steps across weekdays and weekends
```{r}
avgStepsInWeek <- aggregate(steps ~ interval + dayType, data=activityFilled, mean)

library(ggplot2)
ggplot(avgStepsInWeek, aes(interval, steps)) +      
  geom_line() + 
  facet_grid(dayType ~ .) 

```










